---
title: "Lab11"
output: html_document
---

```{r setup, cache = F, include=FALSE}
knitr::opts_chunk$set(error = TRUE, echo = TRUE)
```

I had some trouble with this lab, things were working fine and I came back to make sure I had finished the lab and after I reopened the .rmd file I cannot get some of the commands to be recognized or load. I was able to see the charts before, I'm not sure if I deleted something and that is why it isn't working? 
I don't know, I tried to find ways to work around it or rearrange/reload/update libraries and it didn't look like anything changed.

# BiocManager and Libraries
```{r}
BiocManager::install("phyloseq")
BiocManager::install("dada2")
BiocManager::install("DECIPHER")

library(dada2); packageVersion("dada2")
library(DECIPHER); packageVersion("DECIPHER")
library("phyloseq")
```

```{r}
path <- "data/MiSeq_SOP"
list.files

taxa <- assignTaxonomy(seqtab.nochim, "silva_nr99_v138_train_set.fa.gz", multithread = TRUE)

dna <- DNAStringSet(getSequences('seqtab.nochim'))
load("data/SILVA_SSU_r138_2019.RData")
```


------
# Inspect read quality profiles
```{r}
path <- "data/MiSeq_SOP"

fnFs <- sort(list.files(path, pattern="_R1_001.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_001.fastq", full.names = TRUE))

sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)

plotQualityProfile(fnFs[1:2])

plotQualityProfile(fnRs[1:2])
```

# Filter and Trim
```{r}
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))

names(filtFs) <- sample.names
names(filtRs) <- sample.names

out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(240,160), maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,compress=TRUE, multithread=TRUE) 
head(out)
```

# Learn the Error Rates
```{r}
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
```

# Sample inference
```{r}
dadaFs <- dada(filtFs, err=errF, multithread=TRUE)
dadaRs <- dada(filtRs, err=errR, multithread=TRUE)
dadaFs[[1]]
```

# Merge paired reads
```{r}
mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
```

# Construct the sequence table
```{r}
seqtab <- makeSequenceTable(mergers)
dim(seqtab)

table(nchar(getSequences(seqtab)))
```

# Remove chimeras
```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)

sum(seqtab.nochim)/sum(seqtab)
```

# Track reads through the pipeline
```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))
# If processing a single sample, remove the sapply calls: e.g. replace sapply(dadaFs, getN) with getN(dadaFs)
colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
head(track)
```

# Assign taxonomy
```{r}
taxa <- assignTaxonomy('seqtab.nochim', "~/tax/silva_nr_v132_train_set.fa.gz", multithread=TRUE)

taxa.print <- taxa # Removing sequence rownames for display only
rownames(taxa.print) <- NULL
head(taxa.print)
```

# DECIPHER
```{r}
library(DECIPHER); packageVersion("DECIPHER")

dna <- DNAStringSet(getSequences(seqtab.nochim)) # Create a DNAStringSet from the ASVs
load("~/tax/IDTaxa/SILVA_SSU_r132_March2018.RData") # CHANGE TO THE PATH OF YOUR TRAINING SET
ids <- IdTaxa(dna, trainingSet, strand="top", processors=NULL, verbose=FALSE) # use all processors
ranks <- c("domain", "phylum", "class", "order", "family", "genus", "species") # ranks of interest
# Convert the output object of class "Taxa" to a matrix analogous to the output from assignTaxonomy
taxid <- t(sapply(ids, function(x) {
        m <- match(ranks, x$rank)
        taxa <- x$taxon[m]
        taxa[startsWith(taxa, "unclassified_")] <- NA
        taxa
}))
colnames(taxid) <- ranks; rownames(taxid) <- getSequences(seqtab.nochim)
```

# Evaluate accuracy
```{r}
unqs.mock <- seqtab.nochim["Mock",]
unqs.mock <- sort(unqs.mock[unqs.mock>0], decreasing=TRUE) # Drop ASVs absent in the Mock
cat("DADA2 inferred", length(unqs.mock), "sample sequences present in the Mock community.\n")

mock.ref <- getSequences(file.path(path, "HMP_MOCK.v35.fasta"))
match.ref <- sum(sapply(names(unqs.mock), function(x) any(grepl(x, mock.ref))))
cat("Of those,", sum(match.ref), "were exact matches to the expected reference sequences.\n")
```

# Bonus: handoff to phyloseq
```{r}
library(phyloseq); packageVersion("phyloseq")
library(Biostrings); packageVersion("Biostrings")
library(ggplot2); packageVersion("ggplot2")
theme_set(theme_bw())

samples.out <- rownames('seqtab.nochim')
subject <- sapply(strsplit(samples.out, "D"), `[`, 1)
gender <- substr(subject,1,1)
subject <- substr(subject,2,999)
day <- as.integer(sapply(strsplit(samples.out, "D"), `[`, 2))
samdf <- data.frame(Subject=subject, Gender=gender, Day=day)
samdf$When <- "Early"
samdf$When[samdf$Day>100] <- "Late"
rownames(samdf) <- samples.out

ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               sample_data(samdf), 
               tax_table(taxa))
ps <- prune_samples(sample_names(ps), ps)

dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps

plot_richness(ps, x="Day", measures=c("Shannon", "Simpson"), color="When")

ps.prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance="bray")

plot_ordination(ps.prop, ord.nmds.bray, color="When", title="Bray NMDS")

top20 <- names(sort(taxa_sums(ps), decreasing=TRUE))[1:20]
ps.top20 <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20)
plot_bar(ps.top20, x="Day", fill="Family") + facet_wrap(~When, scales="free_x")
```




```


