---
title: "Lab9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Base R Functions:
- grep(), grepl(): Search for matches of a regular expression/pattern in a character vector. grep() returns the indices into the character vector that contain a match or the specific strings that happen to have the match grepl() returns a TRUE/FALSE vector indicating which elements of the character vector contain a match
- nregexpr(), gregexpr(): Search a character vector for regular expression matches and return the indices of the string where the match begins and the length of the match
- sub(), gsub(): Search a character vector for regular expression matches and replace that match with another string
- regexec(): This function searches a character vector for a regular expression, much like regexpr(), but it will additionally return the locations of any parenthesized sub-expressions. Probably easier to explain through demonstration.

The stringr Package:
- str_subset() is much like grep(value = TRUE) and returns a character vector of strings that contain a given match
- str_detect() is essentially equivalent grepl()
- str_extract() plays the role of regexpr() and regmatches(), extracting the matches from the output
 - str_match() does the job of regexec() by provide a matrix containing the parenthesized sub-expressions

## Matching patterns with regular expressions

#libraries
```{r}
library(stringr)
library(tidyverse)
```

# Basic matches
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")

str_view(x, ".a.")

dot <- "\\."
writeLines(dot)

x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")

x <- "a\\b"
writeLines(x)

str_view(x, "\\\\")
```

# 14.3.1.1 Exercise
1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".
    "\" functions to escape the next character. "\\" resolves \ which will then escape the next character. In "\\\" the first two \\ will resolve to a \ which then escapes the next character.

# 14.3.2 Anchors
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
str_view(x, "a$")

x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")

str_view(x, "^apple$")
```

# 14.3.2.1 Exercise
1. How would you match the literal string "$^$"?
```{r}
str_view(c("$^$", "ab$^$sfas"), "^\\$\\^\\$$", match = TRUE)
```

# 14.3.3 Character classes and alternatives
- \d : matches any digit.
- \s : matches any whitespace (e.g. space, tab, newline).
- [abc] : matches a, b, or c.
- [^abc] : matches anything except a, b, or c.

```{r}
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")

str_view(x, ".[*]c")

str_view(x, "a[ ]")

x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

# 14.3.3.1 Exercise
```{r}
str_subset(stringr::words, "^[aeiou]")

str_view(stringr::words, "[aeiou]", match = FALSE)

str_subset(stringr::words, "[^e]ed$")

str_subset(stringr::words, "i(ng|se)$")
```

# 14.3.4 Repetition
- ?: 0 or 1
- +: 1 or more
- *: 0 or more

- {n}: exactly n
- {n,}: n or more
- {,m}: at most m
- {n,m}: between n and m
```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")

str_view(x, "CC+")

str_view(x, 'C[LX]+')

str_view(x, "C{2}")

str_view(x, "C{2,}")

str_view(x, "C{2,}")

str_view(x, 'C{2,3}?')

str_view(x, 'C[LX]+?')
```

# 14.3.4.1 Exercise
1. Describe the equivalents of ?, +, * in {n,m} form.
    In {m,n} form...
    ? matches between 0 and 1, a short string
    + matches more than 1
    * matches more than 0

# 14.3.5 Grouping and backreferences
```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

# 14.3.5.1 Exercise
1. Describe, in words, what these expressions will match:
- (.)\1\1 matches the same character three times in a row, ex. aaa
- "(.)(.)\\2\\1" matches two characters first and the same pattern of two characters but reversed, ex. abba
- (..)\1 matches repeated pairs of characters, ex. abab
- "(.).\\1.\\1" I'm not sure about the last two...
- "(.)(.)(.).*\\3\\2\\1" 


# 14.4.1 Detect matches
```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
sum(str_detect(words, "^t"))
mean(str_detect(words, "[aeiou]$"))

no_vowels_1 <- !str_detect(words, "[aeiou]")
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)

words[str_detect(words, "x$")]
str_subset(words, "x$")

df <- tibble(
  word = words, 
  i = seq_along(word))

df %>% 
  filter(str_detect(word, "x$"))

x <- c("apple", "banana", "pear")
str_count(x, "a")
mean(str_count(words, "[aeiou]"))

df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]"))

str_count("abababa", "aba")
str_view_all("abababa", "aba")
```

# 14.4.1.1 Exercise
1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
--> Find all words that start or end with x.
```{r}
words[str_detect(words, "^x|x$")]

# or

start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]
```
--> Find all words that start with a vowel and end with a consonant.
```{r}
str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()

# or

start_with_vowel <- str_detect(words, "^[aeiou]")
end_with_consonant <- str_detect(words, "[^aeiou]$")
words[start_with_vowel & end_with_consonant] %>% head()
```
--> Are there any words that contain at least one of each different vowel?
```{r}
words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```

# 14.4.2 Extract matches
```{r}
length(sentences)
head(sentences)

colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match

has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)

has_colour

more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)

str_extract(more, colour_match)
str_extract_all(more, colour_match)
str_extract_all(more, colour_match, simplify = TRUE)

x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

# 14.4.2.1 Exercise
1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.
```{r}
colour_matched <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_matched, match = TRUE)
```

# 14.4.3 Grouped matches
```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)

has_noun %>% 
  str_match(noun)

tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE)
```

# 14.4.3.1 Exercise
1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
```{r}
wordafternum <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"

sentences[str_detect(sentences, wordafternum)] %>%
  str_extract(wordafternum)
```

# 14.4.4 Replacing matches
```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")

str_replace_all(x, "[aeiou]", "-")

x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))

sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

# 14.4.4.1 Exercise
1. Replace all forward slashes in a string with backslashes.
```{r}
str_replace_all("past/present/future", "/", "\\\\")
```

# 14.4.5 Splitting
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")

"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]

sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)

fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)

x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))

str_split(x, " ")[[1]]
str_split(x, boundary("word"))[[1]]
```

# 14.4.5.1 Exercise
1. Split up a string like “apples, pears, and bananas” into individual components.
```{r}
z <- c("apples, pears, and bananas")
str_split(z, ", +(and +)?")[[1]]
```
